---
title: "BOX UHS Download"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
```

```{r}
library(tidyverse)
kprint <- function(x) {
  if(interactive()) {
    print(x)
  } else {
    knitr::kable(x)
  }
}
```

## Useful Columns for Modelers

The following columns would be useful to share with campus modelers. They appear to contain no PHI or PII but have sufficient data for modelers to study patterns of infection. Please note that this is based on data through 5 October 2020. See details below for study of fields

- RANDID (carefully randomized from PatientID)
- Race
- ResultDate, CollectionDate, ConfirmedDate, TracedDate (dropping time of day)
- QuarantineStatus, HasSymptoms, NumberOfContacts
- TestLocation, LabName, Result
- Hall_Name
- Neighborhood (LakeShore or SouthEast)
- RANDHOUS (new randomization within hall and across houses and floors)
- RANDROOM (room number modulus 100 to remove floor)
- Positive

The following are removed from the data but kept locally in lookup table:

- House_Name (recoded with Floor as RANDROOM)
- Floor (100s number from Room)
- Room (recoded as RANDID using FLIP)

## Access Original Data

```{r eval=FALSE}
From: "wiscit@doit.wisc.edu" <wiscit@doit.wisc.edu>
Date: Tuesday, October 13, 2020 at 1:13 PM

The Box Client ID and Client Secret identify the Box-app that you setup to connect to your Box account. This something that you would need to create for yourself via the Box Dev console:
 
https://uwmadison.app.box.com/developers/console
 
The UW-Madison Box team does not offer any custom Box development or consulting services but here are some links that might help you get started:
 
https://developer.box.com/guides/authentication/select/
 
https://developer.box.com/guides/applications/custom-apps/
 
Please keep in mind that the only authorization type that is supported with UW-Madison Box is Oauth 2.0 authentication:
 
https://kb.wisc.edu/103326
 
Finally, be aware that we can no longer offer unlimited storage with the Box service. Beginning on 6/15/2020 all accounts have a storage quota set. For questions regarding that change please see our Contract & Quota FAQ: 
 
https://kb.wisc.edu/102615
 
Please let me know if you have additional questions. 
 
Regards,
 
Jeannine 
```

```{r eval=FALSE}
library(boxr)
# Ideally use boxr package to box_read() file. However
# Need Client Secret, which I don't yet understand. See email above.
box_auth("byandell@wisc.edu")
filename <- scan("data/housing_data_url.txt", "text", n = 1)
cdc0 <- box_read(filename)
```

```{r eval=FALSE}
# This sometimes work, but generally gives empty file if done in Rmarkdown.
# What works is to download file and use that directly.
filename <- scan("data/housing_data_url.txt", "text", n = 1)
cdc0 <- read.csv(filename, row.names = NULL)
```

Read data, select columns, do initial cleaning. Focus on students on campus.

```{r}
#filename <- "data/MasterQuery_WithHousing_20201120.csv"
#filename <- "data/MasterQuery_HousingSubset_20201120.csv"
filename <- "data/MasterQuery_HousingSubset_20201202.csv"
cdc0 <- read.csv(filename, row.names = NULL) %>%
  select(PatientID, Group, hall_housingfile, House_Name, Room, Race,
         ResultDate, CollectionDate, ConfirmedDate, TracedDate,
         TestLocation, LabName, Result,
         HasSymptoms, QuarantineStatus, NumberOfContacts) %>%
  rename(Hall_Name = "hall_housingfile") %>%
  filter(!is.na(PatientID),
         Result %in% c("Positive","Negative")) %>%
  mutate(Race = ifelse(Race %in% c("", "DECLINED", "OTHER RACE"), NA, Race),
         Positive = 1 * (Result == "Positive")) %>%
  filter(Hall_Name != "Off campus", Group == "Student")
```

Replace `PatientID` with randomized `RANDID`

```{r}
IDs <- (cdc0 %>% distinct(PatientID) %>% arrange(PatientID))$PatientID
set.seed(1)
Rand <- order(runif(length(IDs)))
cdc0 <- cdc0 %>%
  mutate(RANDID = Rand[match(PatientID, IDs)]) %>%
  select(-PatientID, -Group)
```

Assign `Hall_Name`s to `Neighborhood`.

```{r}
lakeshore <- c("Adams", "Bradley", "Cole", "Dejope", "Kronshage", "Leopold", "Phillips", "Slichter", "Sullivan", "Tripp", "Waters")
southeast <- c("Barnard", "Chadbourne", "Davis", "Merit", "Ogg", "Sellery", "Smith", "Witte")
cdc0 <- cdc0 %>%
  mutate(Neighborhood = "LakeShore",
         Neighborhood = ifelse(Hall_Name %in% southeast, "SouthEast", Neighborhood))
```

Randomize `House_Name` and `Floor` to `RANDHOUS` and `Room` to `RANDROOM`.

```{r}
# Re-randomize House as RANDHOUS within RANDHALL
Rehouse <- cdc0 %>%
  # Break up houses with multiple floors
  mutate(Floor = floor(Room / 100)) %>%
  distinct(Hall_Name, House_Name, Floor) %>%
  group_by(Hall_Name) %>%
  mutate(Floor = Floor,
         RANDHOUS = order(runif(n()))) %>%
  ungroup
cdc0 <- 
  left_join(
    cdc0 %>%
      mutate(Floor = floor(Room / 100),
             RANDROOM = Room %% 100), # Could flip a coin on this.
    Rehouse,
    by = c("Hall_Name", "House_Name", "Floor"))
```

Flip room numbers randomly.

```{r}
cdc0 <- cdc0 %>%
  group_by(Hall_Name, RANDHOUS) %>%
  mutate(Flip = sign(runif(1) - 0.5),
         RANDROOM = ifelse(Flip < 0, 
                           1 + max(RANDROOM) - Flip * RANDROOM, # Flip room number if negative
                           RANDROOM)) %>%
  ungroup
```

Write out codebook.

```{r}
write.csv(cdc0 %>%
  distinct(Neighborhood, Hall_Name, House_Name, RANDHOUS, Floor, Flip) %>%
  arrange(Neighborhood, Hall_Name, House_Name, Floor),
  "data/uhs_codebook.csv", row.names = FALSE)
```

Pivot to have one event per row per student.

```{r}
# Pivot longer; create RANDROOM, filter out missing Date.
cdc <- cdc0 %>%
  select(-House_Name, -Room, -Result, -Floor, -Flip) %>%
  pivot_longer(ResultDate:TracedDate, names_to = "Event", values_to = "Date") %>%
  mutate(Event = factor(Event, c("CollectionDate", "ResultDate",
                                 "ConfirmedDate", "TracedDate")),
         # Remove time of day from Date
         Date = as.Date(Date, "%m/%d/%Y")) %>%
  # For now, remove data with weird dates
  filter(!is.na(Date),
         Date > as.Date("2020-01-01")) %>%
  # Filter out Confirmed or Traced Date that have Positive = 0
  filter(!(Event %in% c("ConfirmedDate", "TracedDate")) | Positive == 1) %>%
# This step is tricky.
# ConfirmedDate and TracedDate are repeated sometimes with old CollectionDate.
  distinct(RANDID, Date, Event, .keep_all = TRUE) %>%
  arrange(Hall_Name, RANDID, Date) %>%
  select(RANDID, RANDROOM, RANDHOUS, Hall_Name, Neighborhood, Date, Event, Positive, everything())
```

```{r}
ggplot(bind_rows(
    cdc %>%
      filter(Event == "CollectionDate") %>%
      count(RANDID, Positive, name = "collections") %>% 
      count(Positive, collections, name = "frequency") %>%
      mutate(Positive = c("Negative","Positive")[1 + Positive]),
    cdc %>%
      filter(Event == "CollectionDate") %>%
      count(RANDID, name = "collections") %>% 
      count(collections, name = "frequency") %>%
      mutate(Positive = "Combined"))) +
  aes(collections, frequency, col = Positive, group = Positive) +
  geom_point() +
  geom_line() +
  ggtitle("Frequency of collections across students")
```

```{r}
write.csv(cdc, "data/uhs_rand.csv", row.names = FALSE)
```

This document summarizes housing information. The R code file can be found in the [pandemic github repository](https://github.com/UW-Madison-DataScience/pandemic/blob/master/box_uhs.Rmd).
