---
title: "CDC_Housing"
author: "`r paste('Brian Yandell,', format(Sys.time(), '%d %B %Y'))`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 7)
```

```{r}
library(tidyverse)
kprint <- function(x) {
  if(interactive()) {
    print(x)
  } else {
    knitr::kable(x)
  }
}
```

```{r}
cdc <- read.csv("data/matchforCDC-PHI.csv") %>%
  select(PatientID, 
         ResultDate, CollectionDate, ConfirmedDate, TracedDate,
         Residence_Hall, fakeroom, FAKEHALL, FAKEHOUS, Positive) %>%
  filter(!is.na(FAKEHALL),
         !is.na(PatientID)) %>%
  mutate(Campus = ifelse(Residence_Hall == "off campus", "off", "on")) %>%
  select(-Residence_Hall)
# Only a few off campus and no positives, so drop them.
cdc <- cdc %>%
  filter(Campus == "on") %>%
  select(-Campus)
```

```{r}
IDs <- sort(unique(cdc$PatientID))
pos <- (cdc %>% filter(Positive == 1))$PatientID
```

```{r}
cdc_date <- cdc %>%
  pivot_longer(ResultDate:TracedDate, names_to = "Event", values_to = "Date") %>%
  distinct(PatientID, Date, Event, .keep_all = TRUE) %>%
  mutate(Event = factor(Event, c("CollectionDate", "ResultDate", "ConfirmedDate", "TracedDate")),
         ID = match(PatientID, IDs),
         Date = as.Date(Date, "%m/%d/%Y")) %>%
  filter(Date > as.Date("2020-01-01"))
```

```{r}
# This does not do what I want.
# Want to replace ID with order of date, but only first instance for each PatientID
idf <- function(ID, dat, group) {
  set <- transpose(
    list(s = split(ID, group),
         d = split(dat, group)))
  out <- map(set, function(x) {
    dmin <- sapply(split(x$d, x$s), min)
    out <- 
    names(out) <- x$s
    out
  })
  names(out) <- NULL
  # want to turn list into vector here****
  as_vector(out)[as.character(ID)]
}
```

```{r}
countroom <- function(room) {
  out <- room %>%
    mutate(PatientID = paste0("N", match(PatientID, sort(unique(PatientID)))))
  if(length(unique(out$PatientID)) > 1) {
    out <- out%>%
      pivot_wider(names_from = "PatientID", values_from = "Positive") %>%
      arrange(Date) %>%
      fill(contains("N")) %>%
      pivot_longer(any_of(paste0("N",1:10)), 
                   names_to = "roommate", values_to = "score") %>%
      ungroup %>%
      group_by(Date) %>%
      summarize(score = sum(score, na.rm = TRUE)) %>%
      ungroup
  }
  out
}
```

```{r}
oneroom <- cdc_date %>%
  filter(Event == "CollectionDate") %>%
  group_by(FAKEHALL, FAKEHOUS, fakeroom, PatientID) %>%
  summarize(Date = ifelse(max(Positive) == 1, min(Date[Positive == 1]), max(Date)),
            Positive = max(Positive)) %>%
  ungroup %>%
  group_by(FAKEHALL, FAKEHOUS, fakeroom) %>%
  arrange(Date) %>%
  mutate(Positive = cumsum(Positive),
         Date = max(Date) - min(Date),
         Positive = max(Positive),
         Roommates = n()) %>%
  ungroup %>%
  filter(Roommates == 2,
         !is.na(Date))
one <- oneroom %>%
  group_by(FAKEHALL, FAKEHOUS, Positive) %>%
  summarize(center = median(Date),
            spread = IQR(Date),
            HouseSize = n()) %>%
  ungroup
```

Number of rooms with 0, 1 or 2 students positive:

```{r}
one %>%
  group_by(FAKEHALL, Positive) %>%
  summarize(HouseSize = sum(HouseSize)) %>%
  ungroup %>%
  mutate(Positive = paste0("Pos", Positive)) %>%
  pivot_wider(names_from = "Positive", values_from = "HouseSize", values_fill = 0) %>%
  mutate(total = Pos0 + Pos1 + Pos2) %>%
  arrange(desc(total)) %>%
  kprint()
```
Next plot summarizes by House (point) within Hall (facet) differences within rooms of last collection date; center = median, spread = IQR.
Subsequent plots show individual students per horizontal line (sometimes with a little jitter), with roommates adjacent to each other. Coloring highlights time course of dates, positive vs negative, or room.

```{r}
ggplot(one %>%
         mutate(Positive = factor(Positive))) +
  aes(center, spread, col = Positive, size = HouseSize) +
  facet_wrap(~ FAKEHALL) +
  geom_jitter(width = 0.1, alpha = 0.25) +
  ggtitle("Days between Last Collection Dates per Room by House within Hall")
```

```{r}
both <- oneroom %>%
  filter(Positive == 2) %>%
  select(FAKEHALL, FAKEHOUS, fakeroom, PatientID)
half <- oneroom %>%
  filter(Positive == 1) %>%
  select(FAKEHALL, FAKEHOUS, fakeroom, PatientID)
cdcID <- cdc_date %>% 
  filter(paste(fakeroom, FAKEHALL, FAKEHOUS) %in% paste(both$fakeroom, both$FAKEHALL, both$FAKEHOUS)) %>%
  group_by(FAKEHALL, FAKEHOUS, PatientID) %>%
  summarize(Date = min(Date),
            fakeroom = fakeroom[1]) %>%
  ungroup %>%
  group_by(FAKEHALL) %>%
  mutate(ID = 0.25 + rank(10000 * FAKEHOUS + fakeroom, ties.method = "random") / 2,
         roomID = rank(10000 * FAKEHOUS + fakeroom)) %>%
  ungroup %>%
  select(FAKEHALL, PatientID, ID, roomID)
cdcIDh <- cdc_date %>% 
  filter(paste(fakeroom, FAKEHALL, FAKEHOUS) %in% paste(half$fakeroom, half$FAKEHALL, half$FAKEHOUS)) %>%
  group_by(FAKEHALL, FAKEHOUS, PatientID) %>%
  summarize(Date = min(Date),
            fakeroom = fakeroom[1]) %>%
  ungroup %>%
  group_by(FAKEHALL) %>%
  mutate(ID = 0.25 +rank(10000 * FAKEHOUS + fakeroom, ties.method = "random") / 2,
         roomID = rank(10000 * FAKEHOUS + fakeroom)) %>%
  ungroup %>%
  select(FAKEHALL, PatientID, ID, roomID)
```

```{r fig.width = 10}
ggplot(
  left_join(
    cdc_date %>% 
      filter(paste(fakeroom, FAKEHALL, FAKEHOUS) %in% paste(both$fakeroom, both$FAKEHALL, both$FAKEHOUS)) %>%
      select(-ID),
    cdcID,
    by = c("FAKEHALL", "PatientID"))) +
  aes(Date, ID, col = Event, group = ID) +
  geom_jitter(width = 0.05) +
  geom_path() +
  facet_wrap(~ FAKEHALL, scales = "free_y") +
  ggtitle("Time Course by Individual when Both Roommates Positive")
```

```{r fig.width = 10}
tmp <- left_join(
    cdc_date %>% 
      filter(paste(fakeroom, FAKEHALL, FAKEHOUS) %in% paste(both$fakeroom, both$FAKEHALL, both$FAKEHOUS),
             Event == "CollectionDate") %>%
      select(-ID),
    cdcID,
    by = c("FAKEHALL", "PatientID"))
ggplot(tmp) +
  aes(Date, ID, col = factor(roomID), group = ID) +
  geom_point() +
  geom_path() +
  facet_wrap(~ FAKEHALL, scales = "free_y") +
  theme(legend.position = "none") +
  scale_color_manual(values = rep(2:7, length.out = length(unique(tmp$roomID)))) +
  ggtitle("Collection Dates by Individual when Both Roommates Positive")
```

```{r}
half_date <- left_join(
    cdc_date %>% 
      filter(paste(fakeroom, FAKEHALL, FAKEHOUS) %in% paste(half$fakeroom, half$FAKEHALL, half$FAKEHOUS),
             Event == "CollectionDate") %>%
      select(-ID),
    cdcIDh,
    by = c("FAKEHALL", "PatientID")) %>%
  group_by(PatientID) %>%
  mutate(Positive = c("Negative","Positive")[1 + max(Positive)]) %>%
  ungroup
```


```{r fig.width = 10}
ggplot(half_date) +
  aes(Date, ID, col = Positive, group = ID) +
  geom_point() +
  geom_path() +
  facet_wrap(~ FAKEHALL, scales = "free_y") +
  ggtitle("Collection Dates by Individual when One Roommate is Positive")
```

```{r}
ggplot(
  left_join(
    cdc_date %>% 
      filter(paste(fakeroom, FAKEHALL, FAKEHOUS) %in% paste(both$fakeroom, both$FAKEHALL, both$FAKEHOUS)) %>%
      select(-ID),
    cdcID,
    by = c("FAKEHALL", "PatientID")) %>%
    filter(FAKEHALL == 5, fakeroom != 14933)) + # not sure why this shows up
  aes(Date, ID, col = Event, group = ID) +
  geom_jitter(width = 0.05, size = 2) +
  facet_wrap(~ fakeroom, scales = "free_y") +
  ggtitle("All Dates for Roommates Both Positive in Fake Hall 5")
```

```{r fig.width = 10}
ggplot(
  left_join(
    cdc_date %>% 
      filter(paste(fakeroom, FAKEHALL, FAKEHOUS) %in% paste(both$fakeroom, both$FAKEHALL, both$FAKEHOUS),
             Event == "CollectionDate") %>%
      select(-ID),
    cdcID,
    by = c("FAKEHALL", "PatientID")) %>%
    filter(FAKEHALL == 16)) +
  aes(Date, ID, col = factor(roomID), group = ID) +
  geom_point() +
  geom_path() +
  facet_wrap(~ FAKEHOUS, scales = "free_y") +
  theme(legend.position = "none") +
  scale_color_manual(values = rep(2:7, length.out = length(unique(tmp$roomID)))) +
 ggtitle("Collection Dates for Fake Hall 16 Houses with Roommates Both Positive")
```

```{r}
knitr::knit_exit()
```

### Older Stuff

```{r}
tmp <- (cdc %>% count(PatientID, name = "tests") %>% count(tests) %>% t)
colnames(tmp) <- rep("", ncol(tmp))
tmp
```


```{r}
tmp <- cdc %>% 
  count(FAKEHALL, FAKEHOUS, PatientID, name = "count") %>%
  mutate(count = factor(ifelse(count > 5, ">5", as.character(count)), c(1:5, ">5"))) %>%
  count(FAKEHALL, FAKEHOUS, count) %>%
  mutate(FAKEHALL = factor(FAKEHALL))
ggplot(tmp) +
  aes(count, n, col = FAKEHALL, group = FAKEHALL) +
  geom_point() +
  geom_path() +
  scale_y_log10() +
  theme(legend.position = "none") +
  xlab("Number of Tests") +
  ylab("Student Count") +
  ggtitle("How Many Tests per Student by Hall?")
```

```{r}
ggplot(cdc_date %>% filter(PatientID %in% pos, FAKEHALL == FAKEHALL[1])) +
  aes(Date, ID, col = Event, group = ID) +
  geom_point() +
  geom_path() +
  ggtitle("Positive")
```


```{r}
ggplot(cdc_date %>% filter(PatientID %in% pos, FAKEHALL == FAKEHALL[1], Event %in% c("ConfirmedDate", "TracedDate"))) +
  aes(Date, ID, col = Event, group = ID) +
  geom_point() +
  geom_path() +
  ggtitle("Positive")
```


```{r}
ggplot(cdc_date %>% 
         filter(PatientID %in% pos, 
                FAKEHALL == FAKEHALL[1], 
                Event %in% c("ConfirmedDate", "TracedDate")) %>%
         mutate(ID = idf(ID, Date, FAKEHOUS))) +
  aes(Date, ID, col = Event, group = ID) +
  geom_point() +
  geom_path() +
  facet_wrap(~ FAKEHOUS, scales = "free_y") +
  ggtitle("Positive")
```

Number of roommates positive by house.

```{r}
cdc_date %>%
  filter(PatientID %in% pos, FAKEHALL == FAKEHALL[1], Event %in% c("ConfirmedDate", "TracedDate")) %>%
  count(fakeroom, FAKEHOUS, name = "nc") %>%
  count(FAKEHOUS, nc) %>%
  mutate(nc = paste0("R", nc)) %>%
  pivot_wider(names_from = "nc", values_from = "n", values_fill = 0) %>%
  arrange(desc(R1 + R2 + R3 + R4))
```
Now would like to look at both positive and negative by house and room to see pattern.


```{r}
ggplot(cdc_date %>% filter(FAKEHALL == FAKEHALL[1], Event %in% c("ConfirmedDate")) %>%
         mutate(Outcome = c("Negative","Positive")[1+Positive])) +
  aes(Date, ID, col = Outcome, group = ID) +
  geom_point() +
  geom_path() +
  facet_wrap(~ FAKEHOUS, scales = "free_y") +
  ggtitle("Negative & Positive")
```

