---
title: "Cornell Simulations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

These data are simulation summaries for Cornell model used in <https://data-viz.it.wisc.edu/cornell-parameter-sweep/>.

```{r}
library(tidyverse)
library(RPostgres)
library(odbc)
library(DBI)
library(skimr)
library(broom)
kprint <- function(x) {
  if(interactive()) {
    print(x)
  } else {
    options(knitr.kable.NA = '')
    knitr::kable(x)
  }
} 
```

```{r}
info <- scan("data/cornell.txt", "text")
con <- dbConnect(RPostgres::Postgres(),
                 dbname = "test",
                 host = info[1],
                 port = 5432,
                 user = "postgres",
                 password = info[2])
```


```{r}
# Metrics with sim_id and group_number 
metrics <- dbGetQuery(con, 'Select * from metrics') %>% 
  as_tibble() %>% 
  pivot_longer(cols = -c(sim_id, group_number), 
               names_to = "metric_name", 
               values_to = "metric_value")
```

```{r}
## Parameters for each simulation 
all_params <-
  tbl(con, "group_params") %>%
  distinct() %>%
  collect() %>% 
  mutate(across(where(bit64::is.integer64), as.numeric))

all_params_wide <-
  all_params %>%
  select(-`_scenario_name`) %>%
  pivot_wider(id_cols = sim_id,
              names_from = group_number,
              values_from = c(everything(), -sim_id, -group_number)) %>%
  mutate(across(everything(), ~ as.factor(.x)))
```

```{r}
skim_df <- 
  all_params_wide %>% 
  skimr::skim() %>% 
  filter(factor.n_unique > 1)
```

```{r}
tmpfn <- function(x, y = "", sym = "%", space = " ") {
  paste0(y, space, 100 * signif(as.numeric(as.character(x)), 2), sym)
}
```


```{r}
Groups <- c("UG_on", "UG_off", "Grad_res", "Grad_tea",
                   "Staff_tea", "Staff_res", "Staff_off", "Community")
dat <- metrics %>% 
         mutate(metric_name = make.names(metric_name)) %>%
         pivot_wider(names_from = "metric_name", values_from = "metric_value") %>% 
         left_join(all_params_wide %>% 
                     select(sim_id, skim_df$skim_variable),
                   by = "sim_id") %>%
  rename(contact_UG_on = "contact_rate_multiplier_0",
         contact_UG_off = "contact_rate_multiplier_1",
         contact_Grad_res = "contact_rate_multiplier_2",
         contact_Grad_tea = "contact_rate_multiplier_3",
         contact_Staff = "contact_rate_multiplier_4",
         prev_UG_on = "initial_ID_prevalence_0",
         prev_UG_off = "initial_ID_prevalence_1",
         test_UG_on = "test_population_fraction_0",
         test_UG_off = "test_population_fraction_1",
         Group = "group_number") %>%
  mutate(test_UG = factor(paste0(tmpfn(test_UG_on, "T+", "", ""), tmpfn(test_UG_off, "-", "", "")),
                          c("T+7.1-0.29", "T+7.1-7.1", "T+14-0.29", "T+14-7.1", "T+29-0.29", "T+29-7.1")),
         prev_UG = factor(paste0(tmpfn(prev_UG_on, "P+", "", ""), tmpfn(prev_UG_off, "-", "", "")),
                          c("P+0.29-0.29", "P+0.29-3.3", "P+3.3-0.29", "P+3.3-3.3")),
         test_UG_on = factor(tmpfn(test_UG_on, "T+", "", ""),
                             c("T+7.1", "T+14", "T+29")),
         test_UG_off = factor(tmpfn(test_UG_off, "T-", "", ""),
                              c("T-0.29", "T-7.1")),
         prev_UG_off = factor(tmpfn(prev_UG_off, "P-", "", ""),
                              c("P-0.29", "P-3.3")),
         prev_UG_on = factor(tmpfn(prev_UG_on, "P+", "", ""),
                             c("P+0.29", "P+3.3")),
         Group = factor(Groups[1 + as.integer(Group)], Groups))
```

```{r}
ggplot(dat %>%
         filter(contact_UG_on == 1.25,
                contact_UG_off == 1,
                contact_Grad_res == 1,
                contact_Grad_tea == 1,
                contact_Staff == 1)) +
  aes(Time.of.peak.COVID.19.cases, Peak.active.cases,
      col = Group) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
  scale_y_log10() +
  facet_grid(test_UG ~ prev_UG) +
  ggtitle("Prevalence by Test for all Groups at High Contact Rates")
```

```{r}
ggplot(dat %>%
         filter(contact_UG_off == 1,
                contact_Grad_res == 1,
                contact_Grad_tea == 1,
                contact_Staff == 1)) +
  aes(Time.of.peak.COVID.19.cases, Peak.active.cases,
      col = prev_UG, group = prev_UG) +
  geom_point() +
  geom_line() +
  scale_y_log10() +
  facet_grid(test_UG ~ Group) +
  ggtitle("Prevalence by Test for all Groups over On Campus Contact Rates")
```
Fit overall model with two-factor interactions. Many significant.
Then fit model by group with two-factor interactions.

```{r eval = FALSE}
fits <- lm(log10(Peak.active.cases) ~
             (Group + contact_UG_on + contact_UG_off +
             contact_Grad_res + contact_Grad_tea +
             contact_Staff + 
             test_UG_on + test_UG_off +
             prev_UG_on + prev_UG_off)^2, dat)
fitss <- step(fits, trace = 0)
tidy(anova(fitss))[, c("term","p.value")] %>%
  mutate(p.value = round(p.value, 4),
         p.value = ifelse(p.value > 0.05, NA, p.value)) %>%
  filter(!is.na(p.value))
```

Table of P value * 10^4 if < 0.05, for main effects and 2-way interactions by group.

```{r}
fitfn <- function(dat) {
  fits <- lm(log10(Peak.active.cases) ~
             (contact_UG_on + contact_UG_off +
             contact_Grad_res + contact_Grad_tea +
             contact_Staff + 
             test_UG + prev_UG)^2, dat)
  step(fits, trace = 0) 
}
fit_group <- map(
  split(dat, dat$Group),
  fitfn)
(aov_group <- bind_rows(
  lapply(fit_group, function(x) tidy(anova(x))[,c("term","p.value")]),
  .id = "Group") %>%
  mutate(p.value = round(p.value, 4),
         p.value = ifelse(p.value > 0.05, NA, p.value * 10000)) %>%
  pivot_wider(names_from = "Group", values_from = "p.value") %>%
  filter(term != "Residuals")) %>%
  kprint()
```
```{r}
group_plot <- function(dat, group_name, group_title) {
  ggplot(dat %>%
              filter(Group == group_name)) +
    aes(contact_UG_on, Peak.active.cases,
        col = test_UG_on, group = test_UG_on) +
    geom_jitter(height = 0, width = 0.25) +
    geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
    scale_y_log10() +
    facet_grid(test_UG_off ~ prev_UG, scales = "free") +
    ylab("Peak Active Cases") +
    ggtitle(group_title)
}
```

```{r}
group_plot(dat, "UG_on", "UG On Campus")
```

```{r}
group_plot(dat, "UG_off", "UG Off Campus")
```

```{r}
group_plot(dat, "Grad_res", "Grad Researchers")
```

```{r}
group_plot(dat, "Grad_tea", "Grad Instructors")
```

```{r}
group_plot(dat, "Staff_tea", "Staff Instructors")
```

```{r}
group_plot(dat, "Staff_res", "Staff Researchers")
```

```{r}
group_plot(dat, "Staff_off", "Staff Off Campus")
```

```{r}
group_plot(dat, "Community", "Madison Community")
```

#### Some investigation of Grad Instuctors Key Interactions

It seems that `Grad_tea` and `Staff_tea` are different and simpler in their model, and that `contact_Grad_tea` interactions seem modest. Use fits to suggest interesting plots.

```{r}
ggplot(dat %>%
         filter(Group == "Grad_tea",
                contact_Grad_res == 1,
                contact_Grad_tea == 1,
                contact_Staff == 1)) +
  aes(Time.of.peak.COVID.19.cases, Peak.active.cases,
      col = contact_UG_on, shape = contact_UG_off, group = contact_UG_on) +
  geom_point() +
  geom_path() +
  scale_y_log10() +
  facet_grid(test_UG ~ prev_UG, scales = "free") +
  ggtitle("Graduate Student Instuctors")
```

```{r}
ggplot(dat %>%
         filter(Group == "Grad_tea")) +
  aes(prev_UG, Peak.active.cases,
      col = contact_Staff) +
  geom_boxplot() +
  ggtitle("Graduate Student Instuctors")
```
```{r}
fit <- lm(Peak.active.cases ~ contact_UG_on + contact_UG_off + 
            contact_Grad_tea + test_UG,
          dat %>%
            filter(Group == "Grad_tea"))
datr <- dat %>%
  filter(Group == "Grad_tea") %>%
  mutate(resid = resid(fit) + mean(Peak.active.cases),
         prev_UG_on = str_remove(prev_UG_on, "prev_UG_on "),
         prev_UG_off = str_remove(prev_UG_off, "prev_UG_off "))
```

```{r}
ggplot(datr) +
  aes(prev_UG_on, resid,
      col = prev_UG_off) +
  geom_boxplot() +
  facet_wrap(~ paste0("contact_Staff ", contact_Staff)) +
  ylab("Peak removing Main Effects") +
  ggtitle("Graduate Student Instuctor Interaction")
```

```{r}
ggplot(datr %>%
         group_by(prev_UG_on, prev_UG_off, contact_Staff) %>%
         summarize(resid = mean(resid), .groups = "drop")) +
  aes(prev_UG_on, resid,
      col = prev_UG_off, group = prev_UG_off) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~ paste0("contact_Staff ", contact_Staff)) +
  ylab("Peak removing Main Effects") +
  ggtitle("Graduate Student Instuctor Interaction")

```


