---
title: "Housing Move-In"
author: "`r paste('Brian Yandell,', format(Sys.time(), '%d %B %Y'))`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
kprint <- function(x) {
  if(interactive()) {
    print(x)
  } else {
    knitr::kable(x)
  }
}
```

### Testing for all Residents

```{r}
move_all <- read_csv("data/Housing Students Test Results_09012020.csv", 
                 col_types = cols(QuarantineNotes = col_character())) %>%
  rename(TracedDate = "Traced Date",
         NumberContacts = "Number of Contacts",
         ResidenceHall = "Residence Hall",
         AdmitType = "Admit Type") %>%
  mutate(ResultDate = as.POSIXct(ResultDate, format = "%m/%d/%y"),
         TracedDate = as.POSIXct(TracedDate, format = "%m/%d/%y"),
         SymptomaticDate = as.POSIXct(SymptomaticDate, format = "%m/%d/%y"),
         Result = ifelse(Result %in% c("Negative", "Positive"), Result, "Negative"))
nstud <- nrow(distinct(move_all, NEWID))
```

This includes testing before move-in week as well as testing during that period
between `r min(move_all$ResultDate)` and `r max(move_all$ResultDate)`.

There are `r nstud` students reported in housing with the following distribution of repeats:

```{r}
move_all %>%
  count(NEWID, name = "repeats") %>%
  count(repeats) %>%
  kprint()
```

```{r}
tmp <- move_all %>% 
  filter(Result == "Positive") %>%
  count(NEWID, name = "repeats") %>% 
  count(repeats, name = "Positive")
```

There were `r sum(tmp$Positive)` students (`r round(100 * sum(tmp$Positive) / nstud, 2)`%) who tested positive with the following frequency:

```{r}
tmp %>%
  kprint()
```

Here is the breakdown of repeats by number of positive results per student:

```{r}
(tmp <- move_all %>% 
  count(NEWID, ResidenceHall, Result) %>%
  pivot_wider(names_from = "Result", values_from = "n", values_fill = 0) %>%
  count(Negative, Positive) %>%
  rename(repeats = "Negative") %>%
  mutate(repeats = Positive + repeats,
         Positive = paste0(Positive, "_Positive")) %>%
  pivot_wider(names_from = "Positive", values_from = "n", values_fill = 0)) %>%
  kprint()
```

Thus, there were `r sum(tmp[-1,-1])` students (`r round(100* sum(tmp[-1,-1]) / sum(tmp[,-1]), 2)`%) who had repeat tests, with `r sum(tmp[-1, "0_Positive"])` of those (`r round(100* sum(tmp[-1,"0_Positive"]) / sum(tmp[,-1]), 2)`%) having no positive results. **With this small a percentage of repeats, they do not substantially affect results.**

### Testing only during Move-In Week

```{r}
move <- move_all %>%
  filter(ResultDate >= as.POSIXct("2020-08-25") & ResultDate <= as.POSIXct("2020-09-01"))
nstud_in <- nrow(distinct(move, NEWID))
```

Now we restrict attention to student tests during move-in week,
between `r min(move$ResultDate)` and `r max(move$ResultDate)`.

Of the `r nstud` students with records in residence halls, `r nstud_in` got tested during move-in (`r round(100 * nstud_in / nstud, 2)`%). They have the following distribution of repeats during move-in week:

```{r}
move %>%
  count(NEWID, name = "repeats") %>%
  count(repeats) %>%
  kprint()
```

```{r}
tmp <- move %>% 
  filter(Result == "Positive") %>%
  count(NEWID, name = "repeats") %>% 
  count(repeats, name = "Positive")
```

There were `r sum(tmp$Positive)` students (`r round(100 * sum(tmp$Positive) / nrow(distinct(move, NEWID)), 2)`%) who tested positive with the following frequency:

```{r}
tmp %>%
  kprint()
```

Here is the breakdown of repeats by number of positive results per student:

```{r}
(tmp <- move %>% 
  count(NEWID, ResidenceHall, Result) %>%
  pivot_wider(names_from = "Result", values_from = "n", values_fill = 0) %>%
  count(Negative, Positive) %>%
  rename(repeats = "Negative") %>%
  mutate(repeats = Positive + repeats,
         Positive = paste0(Positive, "_Positive")) %>%
  pivot_wider(names_from = "Positive", values_from = "n", values_fill = 0)) %>%
  kprint()
```

Thus, there were `r sum(tmp[-1,-1])` students (`r round(100* sum(tmp[-1,-1]) / sum(tmp[,-1]), 2)`%) who had repeat tests, with `r sum(tmp[-1, "0_Positive"])` of those (`r round(100* sum(tmp[-1,"0_Positive"]) / sum(tmp[,-1]), 2)`%) having no positive results. **With this small a percentage of repeats, they do not substantially affect results.**

The positive results during move-in week were recorded at these locations:

```{r}
move %>%
  filter(Result == "Positive") %>%
  count(LocationName, AdmitType) %>%
  pivot_wider(names_from = "AdmitType", values_from = "n") %>%
  kprint()
```

### Contacts and Residence Halls

Here we return to all students tested between `r min(move_all$ResultDate)` and `r max(move_all$ResultDate)`.
The following table shows how many students provided a given number of contacts, separated by result status:

```{r}
move_all %>%
  group_by(NEWID) %>%
  summarize(NumberContacts = sum(NumberContacts),
            Positive = c("All_Negative", "Positive")[1 + any(Result == "Positive")]) %>%
  ungroup() %>%
  count(Positive, NumberContacts) %>%
  filter(NumberContacts > 0) %>%
  pivot_wider(names_from = "Positive", values_from = "n", values_fill = 0) %>%
  arrange(NumberContacts) %>%
  kprint()
```

The following cases had symptoms identified:

```{r}
move_all %>%
  filter(!is.na(HasSymptoms) & HasSymptoms) %>%
  select(Result, ResidenceHall, QuarantineStatus, SymptomaticDate) %>%
  kprint()
```

While the dorm may not seem relevant to disease status yet, here is the breakdown:

```{r}
move_all %>%
  group_by(NEWID, ResidenceHall) %>%
  summarize(Positive = c("All_Negative", "Positive")[1 + any(Result == "Positive")]) %>%
  ungroup() %>%
  count(ResidenceHall, Positive) %>%
  pivot_wider(names_from = "Positive", values_from = "n", values_fill = 0) %>%
  arrange(desc(Positive), desc(All_Negative)) %>%
  mutate(Pct_Pos_Dorm = round(100 * Positive / (Positive + All_Negative), 2)) %>%
  kprint()
```

### Graphs

```{r}
move_pos <- move_all %>%
  filter(Result == "Positive") %>%
  mutate(PosID = rank(jitter(as.numeric(ResultDate)))) %>%
  select(NEWID, PosID)
```

```{r}
ggplot(
  full_join(move_pos,
            move_all %>% filter(NEWID %in% move_pos$NEWID),
            by = "NEWID")) +
  aes(ResultDate, PosID, col = Result, group = NEWID) +
  geom_point() +
  geom_path() +
  ggtitle("Students with a Positive Test") +
  ylab("")
```

```{r}
move_neg <- move_all %>% 
  filter(!(NEWID %in% move_pos$NEWID)) %>%
  count(NEWID) %>%
  filter(n > 1)
```

```{r}
ggplot(
  move_all %>% 
    filter(NEWID %in% move_neg$NEWID) %>%
    mutate(NEWID = rank(NEWID))) +
  aes(ResultDate, NEWID, col = Result, group = NEWID) +
  geom_point() +
  geom_path()+
  ggtitle("Students with no Positive and multiple Negative Tests") +
  ylab("")
```


