---
title: "Cases and Wastewater"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(RcppRoll)
library(lubridate)
library(patchwork)
```

```{r}
abbrev_site <- function(x) {
  x <- str_remove(x, " W[WP].*")
  x <- str_remove(x, " MSD")
  x <- str_remove(x, " Water &")
  x <- str_remove(x, " Wastewater Utility")
  abbreviate(x, 10)
}
```

```{r}
caseww <- full_join(
  read_csv("data/wastewater/uwm_ww_data_20210507.csv") %>%
    select(wwtp_name, sample_collect_date, capacity_mgd,
           n1_sars_cov2_conc, n2_sars_cov2_conc,
           ppmov_conc, hf183_conc, bcov_rec_rate,
           average_flow_rate, equiv_sewage_amt) %>%
    rename(Date = "sample_collect_date",
           Site = "wwtp_name"),
  read_csv("data/wastewater/uwm_case_data_20210507.csv") %>%
    rename(Date = "SpecCollectedDate",
           Site = "DISPLAY_NA"),
  by = c("Site", "Date")) %>%
  mutate(Site = abbrev_site(Site),
         Day = factor(weekdays(Date), c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")),
         # Week 53 is the last week of the year bleeding into the next year
         Week = epiweek(Date) + 53 * (2021 == year(Date) & epiweek(Date) < 53),
         Week = Week + 1 - min(Week)) %>%
  arrange(Site, Date)
```

```{r}
ww <- caseww %>%
  filter(!is.na(n1_sars_cov2_conc))
case <- caseww %>%
  filter(!is.na(cases))
```


```{r}
rollo <- function(x, n = 7) {
  nn <- min(length(x), n)
  x <- roll_sum(x, nn, na.rm = TRUE)
  if(nn > 1) {
    x <- c(rep(x[1], nn - 1), x)
  }
  x
}
```

Note that wastewater samples are on different days in different sites, and sometimes more than once in a week. Cases are measured most days, somewhat less on Sundays. It is difficult with this distribution of wastewater measurements to determine lead or lag in signal relative to cases.

```{r}
ggplot(ww) +
  aes(Day) +
  geom_bar() +
  facet_wrap(~ Site) +
  theme(axis.text.x = element_text(angle = 45)) +
  ggtitle("Weekday of Wastewater Sample Collection")
```

```{r}
ggplot(ww) +
  aes(Week, Day) +
  geom_jitter(width = 0, height = 0.1, alpha = 0.5) +
  facet_wrap(~ Site) +
  ggtitle("Week and Day of Wastewater Sample Collection")
```

```{r}
ggplot(case) +
  aes(Day) +
  geom_bar() +
  facet_wrap(~ Site) +
  theme(axis.text.x = element_text(angle = 45)) +
  ggtitle("Weekday of Cases")
```

```{r}
ggplot(case) +
  aes(Week, Day) +
  geom_jitter(width = 0, height = 0.1, alpha = 0.5) +
  facet_wrap(~ Site) +
  ggtitle("Week and Day of Cases")
```

```{r}
ggplot(ww) +
  aes(Week,
      n1_sars_cov2_conc) +
  scale_y_log10() +
  geom_line() +
  facet_grid(Site ~ Day) +
  theme(strip.text.y = element_text(angle = 0)) +
  ggtitle("N1 by Week and Day")
```

```{r}
ggplot(ww) +
  aes(Date,
      rollo(n1_sars_cov2_conc)) +
  scale_y_log10() +
  geom_line() +
  facet_wrap(~ Site) +
  ggtitle("N1 Rolling Mean by Date")
```

```{r}
ggplot(case) +
  aes(Week,
      cases) +
  geom_line() +
  facet_grid(Site ~ Day, scales = "free_y") +
  theme(strip.text.y = element_text(angle = 0)) +
  ggtitle("Cases by Week and Day")
```

```{r}
ggplot(case) +
  aes(Week,
      rollo(cases, 7)) +
  geom_line() +
  facet_grid(Site ~ Day, scales = "free_y") +
  theme(strip.text.y = element_text(angle = 0)) +
  ylab("Cases Rolling Week Mean") +
  ggtitle("Rolling Mean of Cases by Week and Day")
```

```{r}
ggplot(case) +
  aes(Date,
      rollo(cases)) +
  geom_line() +
  facet_wrap(~ Site, scales = "free_y") +
   ylab("Cases Rolling Week Mean") +
 ggtitle("Rolling Mean of Cases by Date")
```

## Burlington

Brief look at one site. Note similarity of N1 and N2, odd point for PPMoV, and disparate trend for BCoV.

```{r}
burl <- caseww %>%
  filter(Site == "Burlington") %>%
  select(Week, cases, n1_sars_cov2_conc:bcov_rec_rate) %>%
  group_by(Week) %>%
  summarize(across(everything(), sum, na.rm = TRUE)) %>%
  mutate(across(everything(), function(x) ifelse(x == 0, NA, x)))
```

```{r}
ggplot(
  burl %>%
    pivot_longer(cases:bcov_rec_rate,
                 names_to = "meas", values_to = "values") %>%
    mutate(meas = factor(meas, 
    c("cases", "n1_sars_cov2_conc", "n2_sars_cov2_conc",
      "ppmov_conc", "bcov_rec_rate", "hf183_conc")))) +
  aes(Week, values) +
  geom_path() +
  geom_point() +
  facet_wrap(~ meas, scales = "free_y") +
  ggtitle("Burlington")
```

```{r}
coeff <- 1000
ggplot(burl) +
  aes(Week) +
  geom_path(aes(y = n1_sars_cov2_conc / coeff), col = "red") +
  geom_point(aes(y = n1_sars_cov2_conc / coeff), col = "red") +
  geom_path(aes(y = cases), col = "blue") +
  geom_point(aes(y = cases), col = "blue") +
  scale_y_continuous(name = "Cases",
#                     trans = "log10",
                     sec.axis = sec_axis(~.*coeff, name = "N1")) +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red")
  ) +
  ggtitle("Burlington: Weekly Cases and N1")
```

```{r}
coeff <- 1000
ggplot(caseww %>%
  filter(Site == "Burlington")) +
  aes(Date) +
  geom_path(aes(y = rollo(n1_sars_cov2_conc / coeff)), col = "red") +
  geom_point(aes(y = rollo(n1_sars_cov2_conc / coeff)), col = "red") +
  geom_path(aes(y = rollo(cases)), col = "blue") +
  geom_point(aes(y = rollo(cases)), col = "blue") +
  scale_y_continuous(name = "Cases",
#                     trans = "log10",
                     sec.axis = sec_axis(~.*coeff, name = "N1")) +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red")
  ) +
  ggtitle("Burlington: Rolling Week Sum of Cases and N1")
```

## All Sites by Week

In these plots, N1 is rescaled to 1000 times the cases, which works for some less populated areas, but not for Milwaukee, Green Bay or Racine. Not that there is not a consistent scaling, which might reflect population differences as well as other characteristics of the samples in different locations.

```{r fig.width = 7, fig.height = 10}
coeff <- 1000
ggplot(caseww %>%
    select(Site, Week, cases, n1_sars_cov2_conc:bcov_rec_rate) %>%
    group_by(Site, Week) %>%
    summarize(across(everything(), sum, na.rm = TRUE)) %>%
    mutate(across(everything(), function(x) ifelse(x == 0, NA, x)))) +
  aes(Week) +
  geom_path(aes(y = n1_sars_cov2_conc / coeff), col = "red") +
  geom_point(aes(y = n1_sars_cov2_conc / coeff), col = "red") +
  geom_path(aes(y = cases), col = "blue") +
  geom_point(aes(y = cases), col = "blue") +
  scale_y_continuous(name = "Cases",
#                     trans = "log10",
                     sec.axis = sec_axis(~.*coeff, name = "N1")) +
  facet_wrap(~ Site, ncol = 3, scales = "free_y") +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red")
  ) +
  ggtitle("All Sites: Weekly Cases and N1")
```

```{r fig.width = 7, fig.height = 10}
coeff <- 1000
ggplot(caseww %>%
    select(Site, Date, cases, n1_sars_cov2_conc:bcov_rec_rate) %>%
      arrange(Site, Date)) +
  aes(Date) +
  geom_path(aes(y = rollo(n1_sars_cov2_conc / coeff)), col = "red") +
  geom_point(aes(y = rollo(n1_sars_cov2_conc / coeff)), col = "red") +
  geom_path(aes(y = rollo(cases)), col = "blue") +
  geom_point(aes(y = rollo(cases)), col = "blue") +
  scale_y_continuous(name = "Cases",
#                     trans = "log10",
                     sec.axis = sec_axis(~.*coeff, name = "N1")) +
  facet_wrap(~ Site, ncol = 3, scales = "free_y") +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red")
  ) +
  ggtitle("All Sites: Rolling Week Sum of Cases and N1")
```

## All Sites by Week Rescaled

These plots allow one to focus on the shape of the curves, and the time of the maxima. If N1 were a consistent lead indicator of cases, then the red peak should generally be to the left of the blue peak, which it is not.

```{r fig.width = 7, fig.height = 10}
coeff <- 1
ggplot(
  caseww %>%
    select(Site, Week, cases, n1_sars_cov2_conc:bcov_rec_rate) %>%
    group_by(Site, Week) %>%
    summarize(across(everything(), sum, na.rm = TRUE)) %>%
    mutate(across(everything(), function(x) ifelse(x == 0, NA, x))) %>%
    group_by(Site) %>%
    mutate(across(everything(), function(x) x / max(x, na.rm = TRUE)))) +
  aes(Week) +
  geom_path(aes(y = n1_sars_cov2_conc / coeff), col = "red") +
  geom_point(aes(y = n1_sars_cov2_conc / coeff), col = "red") +
  geom_path(aes(y = cases), col = "blue") +
  geom_point(aes(y = cases), col = "blue") +
  scale_y_continuous(name = "Cases",
#                     trans = "log10",
                     sec.axis = sec_axis(~.*coeff,
                                         labels = NULL, name = "N1")) +
  facet_wrap(~ Site, ncol = 3, scales = "free_y") +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red")
  ) +
  ggtitle("All Sites: Weekly Cases and N1")
```

```{r fig.width = 7, fig.height = 10}
coeff <- 1
ggplot(caseww %>%
    select(Site, Date, cases, n1_sars_cov2_conc:bcov_rec_rate) %>%
      arrange(Site, Date) %>%
    group_by(Site) %>%
    mutate(across(-Date, rollo),
           across(-Date, function(x) x / max(x, na.rm = TRUE)))) +
  aes(Date) +
  geom_path(aes(y = n1_sars_cov2_conc), col = "red") +
  geom_point(aes(y = n1_sars_cov2_conc), col = "red") +
  geom_path(aes(y = cases), col = "blue") +
  geom_point(aes(y = cases), col = "blue") +
  scale_y_continuous(name = "Cases",
#                     trans = "log10",
                     sec.axis = sec_axis(~.*coeff,
                                         labels = NULL, name = "N1")) +
  facet_wrap(~ Site, ncol = 3, scales = "free_y") +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red")
  ) +
  ggtitle("All Sites: Rolling Week Sum of Cases and N1")
```
