---
title: "DSI Campus Model Team Summaries `r format(Sys.time(), '%d %B %Y')`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 7, fig.height = 3)
```

```{r}
library(tidyverse)
```

```{r}
paltiel <- read.csv("data/output_parameter_sweep_20200921.csv") %>%
  mutate(frequency.screening = factor(frequency.screening, unique(frequency.screening)))
```

**Paltiel Harvard/Yale:**
Frequent testing yields more isolation, higher peaks, but fewer people infected. Higher transmission (R0, left to right) and larger new infection shocks (size) lead to more and earlier isolations and infections. Initial infections (set at 500) has minor effect on peak occupancy, but more lead to earlier peak and more total infections.

```{r fig.height = 4.25}
ggplot(paltiel %>% filter(time.to.return.fps.from.isolation == 14,
#                          new.infections.per.shock == 10,
                          initial.infected == 500)) +
  aes(Time.of.Peak.Isolation, Peak.Isolation.Occupancy,
      group = frequency.screening, 
      col = frequency.screening,
      size = new.infections.per.shock) +
  facet_grid(. ~ r0) +
  scale_y_log10() +
  geom_point() +
  geom_path(size = 1) +
  theme(legend.title = element_blank()) +
  ggtitle("Time vs Size of Peak Isolation by Frequency & Shock Size")
```

```{r fig.height = 4.25}
ggplot(paltiel %>% filter(time.to.return.fps.from.isolation == 14,
#                          new.infections.per.shock == 10,
                          initial.infected == 500)) +
  aes(Time.of.Peak.Isolation, Total.Infections,
      group = frequency.screening,
      col = frequency.screening,
      size = new.infections.per.shock) +
  facet_grid(. ~ r0) +
  scale_y_log10() +
  geom_point() +
  geom_path(size = 1) +
  theme(legend.title = element_blank()) +
  ggtitle("Time of Peak vs Total Infections by Frequency & Shock Size")
```

```{r eval=FALSE}
Here is a CSV that Sri and I put together based on the Harvard/Yale model of Paltiel et al.  It includes the following parameters:

Parameter Sweep
  initial.infected = c(outer(c(1,5),10^(1:2),"*"),1000)
  r0 = seq(0.5,1.5,0.25)
  new.infections.per.shock = c(outer(c(1,5),10^(0:1),"*"),100)
  frequency.screening = c("Symptoms Only”,
   "Every 4 Weeks”,
 "Every 3 Weeks”,
 "Every 2 Weeks”,
 "Weekly”,
 "Every 3 Days”,
 "Every 2 Days”,
 "Daily”)
  time.to.return.fps.from.isolation = c(1,2,14)

Parameter Input (Default)
  initial.susceptible = 6800,
  initial.infected = 100,
  r0 = 1.5,
  exogenous.shocks = "Yes",
  frequency.exogenous.shocks.per.day = 7,
  new.infections.per.shock = 10,
  days.incubation = 3,
  time.to.recovery = 14,
  per.asymptotics.advancing.to.symptoms = 0.3,
  symptom.case.fatality.ratio = 0.0005,
  frequency.screening = "Every 2 Weeks",
  test.sensitivity = 0.8,
  test.specificity = 0.98,
  test.cost = 25,
  time.to.return.fps.from.isolation = 1,
  confirmatory.test.cost = 100

The output columns provided in the csv are 
  Peak Isolation Occupancy
  Time of Peak Isolation
  Total Persons Tested in 80 days
  Total Confirmatory Tests Performed
  Average Isolation Unit Census
  Average %TP in Isolation
  Total testing cost
  Total Infections

This only took a few minutes to run for the 3000 parameters in the sweep above, so if there are other values of the parameter sweep you would like to see or different default parameters, please let us know and we can re-run.  This will get incorporated into the dashboard (with visualizations) in the next couple of days.
```

### McGee Washington Model

```{r}
mcgee <- read.csv("data/summary_McGee_model.csv")
tmp <- list()
for(i in c("weekly", "ever","biw")) {
  tmp[[i]] <- mcgee[,grep(i, names(mcgee))]
  names(tmp[[i]]) <- str_remove(names(tmp[[i]]), "[a-z]*_")
}
names(tmp)[2:3] <- c("everyday","biweekly")
mcgee <-
  bind_rows(tmp,
            .id = "frequency") %>%
  mutate(frequency = factor(frequency, c("biweekly", "weekly", "everyday")))
rm(i, tmp)
```

```{r eval=FALSE}
Note that this data set has 18 columns, 3 times the following 6:

"    "_time_end = time in days until the end of the outbreak in days
"    "_total_perc_inf  = total percent of infected until the end of the outbreak
"    "_peak_inf = number of people in the peak of infected (not percent)
"    "_time_peak_inf = time in days until the peak of infected
"    "_peak_iso = number of people in the peak of isolated (not percent)
"    "_time_peak_iso = time in days until the peak of isolated

The first word of each column is the type of testing strategy. The difference of testing strategies in the file that I sent is that I am changing the testing cadence first weekly, then everyday, and finally biweekly. Each row corresponds to one run of 400 simulations. This model has a several number of parameters, I am listing them here, most of them are self-explained but if you have any question please let me know. 
```

```{r eval=FALSE}
Parameters

NUM_COHORTS              = 8            
NUM_NODES_PER_COHORT     = 850           # students per cohort
NUM_TEAMS_PER_COHORT     = 10

MEAN_INTRACOHORT_DEGREE  = 6
PCT_CONTACTS_INTERCOHORT = 0.1
N = NUM_NODES_PER_COHORT*NUM_COHORTS = 6800 susceptible
INIT_EXPOSED = 25

latentPeriod_mean, latentPeriod_coeffvar = 3.0, 0.6
symptomaticPeriod_mean, symptomaticPeriod_coeffvar = 4.0, 0.4
onsetToHospitalizationPeriod_mean, onsetToHospitalizationPeriod_coeffvar = 11.0, 0.45
hospitalizationToDischargePeriod_mean, hospitalizationToDischargePeriod_coeffvar = 11.0, 0.45
hospitalizationToDeathPeriod_mean, hospitalizationToDeathPeriod_coeffvar = 10.0, 0.45

PCT_ASYMPTOMATIC = 0.25
PCT_HOSPITALIZED = 0.1
PCT_FATALITY = 0.06
R0_mean     = 2.37, R0_coeffvar = 1.5

P_GLOBALINTXN = 0.3   #30% of interactions being with incidental or casual contacts outside their set of close contacts

INTERVENTION_START_PCT_INFECTED = 0/100        # population disease prevalence that triggers the start of the TTI interventions
AVERAGE_INTRODUCTIONS_PER_DAY   = 1/14          # expected number of new exogenous exposures per day

TESTING_CADENCE                 = 'weekly'      # how often to do testing (other than self-reporting symptomatics who can get
                                                #   tested any day)
PCT_TESTED_PER_DAY              = 2000/N        # max daily test allotment defined as a percent of population size
TEST_FALSENEG_RATE              = 'temporal'    # test false negative rate, will use FN rate that varies with disease time
                                                # specifies the test sensitivity via the false negative rates to be used;
                                                #   numerical value specifies constant false negative rate;
                                                #   "temporal" specifies that temporal, state-based false negative rates
                                                #   are to be used
MAX_PCT_TESTS_FOR_SYMPTOMATICS  = 0.95          # max percent of daily test allotment to use on self-reporting symptomatics
MAX_PCT_TESTS_FOR_TRACES        = 1000/N        # max percent of daily test allotment to use on contact traces
RANDOM_TESTING_DEGREE_BIAS      = 0             # magnitude of degree bias in random selections for testing, none here

PCT_CONTACTS_TO_TRACE           = 0.0           # percentage of primary cases' contacts that are traced
TRACING_LAG                     = 2             # number of cadence testing days between primary tests and tracing tests
                                                # the number of cadence testing days between identification of a positive case
                                                # and the testing of their traced contacts

ISOLATION_LAG_SYMPTOMATIC       = 1             # number of days between onset of symptoms and self-isolation of symptomatics
ISOLATION_LAG_POSITIVE          = 2             # test turn-around time (TAT): number of days between administration of test
                                                #   and isolation of positive cases
ISOLATION_LAG_CONTACT           = 0             # number of days between a contact being traced and that contact self-isolating

TESTING_COMPLIANCE_RATE_SYMPTOMATIC                  = 0.8    
TESTING_COMPLIANCE_RATE_TRACED                       = 0.5
TESTING_COMPLIANCE_RATE_RANDOM                       = 0.5  # Assume student testing is not mandatory

TRACING_COMPLIANCE_RATE                              = 0.5

ISOLATION_COMPLIANCE_RATE_SYMPTOMATIC_INDIVIDUAL     = 0.4
ISOLATION_COMPLIANCE_RATE_SYMPTOMATIC_GROUPMATE      = 0.3
ISOLATION_COMPLIANCE_RATE_POSITIVE_INDIVIDUAL        = 0.9
ISOLATION_COMPLIANCE_RATE_POSITIVE_GROUPMATE         = 0.8  # Isolate teams with a positive member,
                                                            # but suppose 20% of students need to be in-person
ISOLATION_COMPLIANCE_RATE_POSITIVE_CONTACT           = 0.2
ISOLATION_COMPLIANCE_RATE_POSITIVE_CONTACTGROUPMATE  = 0.5
```


```{r}
ggplot(mcgee) +
  aes(time_peak_iso, peak_iso, col = frequency) +
  geom_smooth(size = 2, se = FALSE) +
  geom_point(alpha = 0.2) +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("Time to Peak Isolation vs Peak of Isolations")
```

```{r}
ggplot(mcgee) +
  aes(time_peak_inf, peak_inf, col = frequency) +
  geom_smooth(size = 2, se = FALSE) +
  geom_point(alpha = 0.2) +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("Time to Peak vs Peak Infections")
```

```{r}
ggplot(mcgee) +
  aes(time_peak_inf, total_perc_inf, col = frequency) +
  geom_smooth(size = 2, se = FALSE) +
  geom_point(alpha = 0.2) +
  scale_x_log10() +
#  scale_y_log10() +
  ggtitle("Time to Peak vs Percent Infected")
```

This document summarizes DSI Campus Model Team efforts to date. The R code file can be found in the [pandemic github repository](https://github.com/UW-Madison-DataScience/pandemic/blob/master/model_summary.Rmd).

