---
title: "Wastewater"
author: "`r paste('Brian Yandell,', format(Sys.time(), '%d %B %Y'))`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4)
```

```{r}
library(tidyverse)
library(readxl)
library(broom)
library(RcppRoll)
kprint <- function(...) {
  if(interactive()) {
    print(...)
  } else {
    knitr::kable(...)
  }
}
```

This is my quick analysis of the data you provide today. I show linear (dashed) and smooth (solid) trends with data on geometric (log10) scale. Caveats I think (Martin correct if I am off base):

- GEO uses N2 values that are suspect and being revisited and remeasured in coming weeks.
- Flow adjustment done for MMSD based on measurements, but approximated for UW.
- Visual attempt to look at relationship of PMMoV to GEO.
- Analysis of covariance considers (linear) time and site only for previous 14 (MMSD) or 28 (UW) days.

**Notes:**

- For MMSD, GEO shows a 10+ fold rise starting in late November, while PMMoV shows 3 fold rise. This suggests an increase in COVID signal relative to PMMoV around Thanksgiving time (and following Halloween, football, etc.).
- There appears to be increasing variability in GEO with increasing PMMoV signal, although challenging to separate that from trends.
- Sellery with a single population seems to show clearer signal that LakeShore, which draws from multiple buildings.

```{r}
# Linked to most current version on command line
filename <- "data/wastewater/UW_MMSD.xlsx"
```

```{r}
spanw <- 0.75
```

```{r warning = FALSE, message = FALSE}
tmpfn <- function(N1, N2) {
  AVG <- sqrt(N1 * N2)
  AVG <- ifelse(is.na(N1), N2, AVG)
  ifelse(is.na(N2), N1, AVG)
}
missing_codes <- c("","NA","0","Undetected","Not Detected",
                   "Field Parameters to be filled in", 
                   "Inhibited-to be re-ran", "#DIV/0!")
sheets <- excel_sheets(filename)
water_MMSD <- read_excel(filename,
                    na = missing_codes,
                    skip = 12,
                    col_types = c(rep("text", 4), rep("numeric", 4)),
                    sheet = "MMSD") %>%
  select(-c(3,6,8)) %>%
  rename(Date = "Collection Date",
         Description = "Sample Description",
         QC = "Yes/No",
         GEO = "Gene Copies / 1L...5",
         PMMoV = "Gene Copies / 1L...7") %>%
  mutate(Site = "MMSD")

water_Interceptors <- read_excel(filename,
                    na = missing_codes,
                    skip = 12,
                    col_types = c(rep("text", 4), rep("numeric", 4)),
                    sheet = "MMSD Interceptor") %>%
  select(-c(3,6,8)) %>%
  rename(Date = "Collection Date",
         Description = "Sample Description",
         QC = "Yes/No",
         GEO = "Gene Copies / 1L...5",
         PMMoV = "Gene Copies / 1L...7") %>%
  mutate(Site = str_remove(str_remove(Description, "^.*Interceptor "), "\\.+ +Sample.*")) %>%
  filter(!is.na(Site),
         Site %in% c("P2","P7","P8","P11","P18"))

water_UW_LakeShore <- read_excel(filename,
                    na = missing_codes,
                    skip = 12,
                    col_types = c(rep("text", 4), rep("numeric", 4)),
                    sheet = "UW-D") %>%
  select(-c(3,6,8)) %>%
  rename(Date = "Collection Date",
         Description = "Sample Description",
         QC = "Yes/No",
         GEO = "Gene Copies / 1L...5",
         PMMoV = "Gene Copies / 1L...7") %>%
  mutate(Site = "UW-LakeShore")

water_UW_Sellery <- read_excel(filename,
                    na = missing_codes,
                    skip = 12,
                    col_types = c(rep("text", 4), rep("numeric", 4)),
                    sheet = "UW-S") %>%
  select(-c(3,6,8)) %>%
  rename(Date = "Collection Date",
         Description = "Sample Description",
         QC = "Yes/No",
         GEO = "Gene Copies / 1L...5",
         PMMoV = "Gene Copies / 1L...7") %>%
  mutate(Site = "UW-Sellery")

water <- 
  bind_rows(water_MMSD,
            water_Interceptors,
            water_UW_LakeShore,
            water_UW_Sellery) %>%
  mutate(Date = as.Date(Date, format = "%m.%d.%Y")) %>%
  filter(!is.na(Date),
         Date > as.Date("2020-09-01"))
```

```{r}
waters <- water %>%
  pivot_longer(GEO:PMMoV, names_to = "assay", values_to = "values") %>%
  filter(!is.na(values))
```

```{r}
predfn <- function(x) {
  x$pred <- 10^predict(lm(log10(values) ~ Date + Site, x))
  x
}
addpred <- function(x) {
  bind_rows(
    map(
      split(x, x$assay),
      predfn),
    .id = "assay")
}
```

```{r}
fitfn <- function(x, add = FALSE, ...) {
  if(add) {
    form <- formula(log10(values) ~ Date + Site)
    elems <- 2:3
    test <- "drop1"
  } else {
    form <- formula(log10(values) ~ Date * Site)
    elems <- 3
    test <- "anova"
  }
  fit <- lm(form, x)
  out <- switch(test,
         anova = broom::tidy(anova(fit)[elems,]),
         drop1 = broom::tidy(drop1(fit, fit, test = "F")[elems,])) %>%
    select(term, statistic, p.value) %>%
    mutate(statistic = round(statistic, 2),
           p.value = signif(p.value, 3))
  if(!add) {
    out <- out %>%
      mutate(SD = signif(sigma(fit), 3))
  }
  out
}
sumfn <- function(x, ...) {
  bind_rows(
    map(
      split(x, x$assay),
      fitfn, ...),
    .id = "assay") %>%
    arrange(term, assay)
}
timefit <- function(x) {
  out <- bind_rows(
    sumfn(x, TRUE),
    sumfn(x)) %>%
  select(-statistic)
  SD <- out %>%
    select(assay, SD) %>%
    filter(!is.na(SD))
  out <- out %>%
    select(-SD) %>%
    pivot_wider(names_from = "term", values_from = "p.value")
  full_join(out, SD) %>%
    kprint()
}
```

### All Sites

```{r warning = FALSE}
ggplot(waters) +
  aes(Date, values) +
  geom_smooth(method = "loess", se = FALSE, formula = "y ~ x", col = "darkgray") +
  geom_point(alpha = 0.5) +
  scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90)) +
  facet_grid(assay ~ Site, scales = "free_y")
```

\newpage
\ 

### MMSD Composite

```{r warning = FALSE}
ggplot(waters %>% filter(Site %in% c("MMSD"))) +
  aes(Date, values) +
  geom_smooth(method = "loess", se = FALSE,
              formula = "y ~ x", span = spanw) +
  geom_point(alpha = 0.5, size = 2) +
  scale_y_log10() +
  facet_wrap(~ assay, scales = "free_y") +
  ggtitle("MMSD Composite")
```

```{r warning = FALSE}
ggplot(water %>% filter(Site %in% c("MMSD"))) +
  aes(PMMoV, GEO, col = Date) +
  geom_smooth(method = "lm", se = FALSE,
              formula = "y ~ x", span = spanw, linetype = "dashed") +
  geom_smooth(method = "loess", se = FALSE,
              formula = "y ~ x", span = spanw) +
  geom_point(alpha = 0.5, size = 2) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~ Site, scales = "free") +
  ggtitle("MMSD Composite")
```

\newpage
\ 

### MMSD 5 Sites

```{r warning = FALSE}
ggplot(waters %>% filter(Site %in% c("P2","P7","P8","P11","P18"))) +
  aes(Date, values, col = Site) +
  geom_smooth(method = "loess", se = FALSE,
              formula = "y ~ x", span = spanw) +
  geom_point(alpha = 0.5, size = 2) +
  scale_y_log10() +
  facet_wrap(~ assay, scales = "free_y") +
  ggtitle("MMSD Interceptors")
```

```{r warning = FALSE}
ggplot(water %>% filter(Site %in% c("P2","P7","P8","P11","P18"))) +
  aes(PMMoV, GEO, col = Date) +
  geom_smooth(method = "lm", se = FALSE,
              formula = "y ~ x", span = spanw, linetype = "dashed") +
  geom_smooth(method = "loess", se = FALSE,
              formula = "y ~ x", span = spanw) +
  geom_point(alpha = 0.5, size = 2) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~ Site, scales = "free")+
  ggtitle("MMSD Interceptors")
```

```{r warning = FALSE}
ggplot(water %>% filter(Site %in% c("P2","P7","P8","P11","P18"))) +
  aes(PMMoV, GEO, col = Site) +
  geom_smooth(method = "lm", se = FALSE,
              formula = "y ~ x", span = spanw) +
  geom_point(alpha = 0.5, size = 2) +
  scale_x_log10() +
  scale_y_log10()+
  ggtitle("MMSD Interceptors")
```

\newpage
\ 

### Campus Sites

```{r warning = FALSE}
ggplot(waters %>% filter(str_detect(Site, "UW"))) +
  aes(Date, values, col = Site) +
  geom_smooth(method = "loess", se = FALSE,
              formula = "y ~ x", span = spanw) +
  geom_point(alpha = 0.5, size = 2) +
  scale_y_log10() +
  facet_wrap(~ assay, scales = "free_y") +
  ggtitle("Campus Sites")
```

```{r warning = FALSE}
ggplot(water %>% filter(str_detect(Site, "UW"))) +
  aes(PMMoV, GEO, col = Date) +
  geom_smooth(method = "lm", se = FALSE,
              formula = "y ~ x", span = spanw, linetype = "dashed") +
  geom_smooth(method = "loess", se = FALSE,
              formula = "y ~ x", span = spanw) +
  geom_point(alpha = 0.5, size = 2) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~ Site, scales = "free") +
  ggtitle("Campus Sites")
```

```{r warning = FALSE}
ggplot(water %>% filter(str_detect(Site, "UW"))) +
  aes(PMMoV, GEO, col = Site) +
  geom_smooth(method = "lm", se = FALSE,
              formula = "y ~ x", span = spanw, linetype = "dashed") +
  geom_smooth(method = "loess", se = FALSE,
              formula = "y ~ x", span = spanw) +
  geom_point(alpha = 0.5, size = 2) +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("Campus Sites")
```

## Analyses over time

Tables show p-values (small values are more significant). SD is in log10 units.

MMSD Interceptors Analysis of Covariance for Past 28 Days

```{r}
timefit(waters %>% 
  filter(Site %in% c("P2","P7","P8","P11","P18"),
         Date >= max(Date) - 28))
```

```{r}
fit_cov <- function(waters, sites, window) {
  tmp <- waters %>% 
    filter(Site %in% sites,
           Date >= max(Date) - window) %>%
    pivot_wider(names_from = "assay", values_from = "values")
  fit <- lm(log10(GEO) ~ PMMoV + Date + Site + Date + Site, tmp)
  bind_rows(
    broom::tidy(drop1(fit, fit, test = "F")[2:4,]) %>%
      select(term, p.value) %>%
      mutate(p.value = signif(p.value, 3)),
    broom::tidy(anova(lm(log10(GEO) ~ PMMoV + Date + Site + PMMoV * Site, tmp))[4,]) %>%
      select(term, p.value) %>%
      mutate(p.value = signif(p.value, 3))) %>%
    mutate(assay = "GEO") %>%
    select(assay, everything()) %>%
    pivot_wider(names_from = "term", values_from = "p.value") %>%
    mutate(SD = signif(sigma(fit), 3)) %>%
    kprint()
}
```

```{r}
fit_cov(waters, c("P2","P7","P8","P11","P18"), 14)
```

Campus Sites Analysis of Covariance for Past 28 Days

```{r}
timefit(waters %>% 
          filter(str_detect(Site, "UW"),
                 Date >= max(Date) - 28))
```

```{r}
fit_cov(waters, c("UW-LakeShore","UW-Sellery"), 28)
```

```{r}
knitr::knit_exit()
```

\newpage
\ 

### Conductivity

```{r}
dat <- waters %>% 
  filter(assay %in% c("AVG", "Conductivity")) %>%
  pivot_wider(names_from = "assay", values_from = "values") %>%
  filter(!is.na(AVG), !is.na(Conductivity))
ggplot(dat) +
  aes(Conductivity, AVG) +
  geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
  geom_point(alpha = 0.5) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~ Site, scales = "free")
```

```{r}
fit <- lm(AVG ~ Conductivity * Site, data = dat)
anova(fit)
fit <- lm(AVG ~ Conductivity + Site, data = dat)
drop1(fit, fit, test = "F")
```

\newpage
\ 

## Covid Cases

Includes smooth curves and linear trend of previous four weeks.

```{r}
roll7 <- function(x) {
  n7 <- min(length(x), 7)
  x <- roll_mean(x, n7, na.rm = TRUE)
  if(n7 > 1) {
    x <- c(rep(x[1], n7 - 1), x)
  }
  x
}
cases <- "data/uhs_rand.csv"
cases <- read.csv(cases) %>%
  filter(Positive == 1,
         Event == "CollectionDate") %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(RANDID, Date) %>%
  # Drop multiple positive tests
  distinct(RANDID, .keep_all = TRUE) %>% 
  # Focus on LakeShore and Sellery
  rename(Site = "Neighborhood") %>%
  mutate(Site = paste0("UW-", Site),
         Site = ifelse(Site == "UW-SouthEast" & Hall_Name %in% c("Sellery","Witte"),
                       paste0("UW-", Hall_Name), Site)) %>%
  count(Date, Site, name = "Positive")
```

```{r}
write.csv(cases, file = "data/uhs_housing_cases.csv", row.names = FALSE)
```

```{r}
ggplot(cases) +
  aes(Date, Positive, col = Site, group = Site) +
  geom_point() +
  geom_smooth(aes(y = Positive), method = "lm", se = FALSE, size = 1,
              col = "black",
              data = cases %>%
                filter(Date >= max(Date) - 28)) +
  geom_smooth(method = "loess", se = FALSE, size = 1, span = spanw) +
  scale_y_log10() +
  facet_wrap(~ Site) +
  ggtitle("Covid Cases on Campus")
```

```{r}
case_water <- 
  full_join(
    water %>%
      select(-Comment, -pH, -Total_Flow) %>%
      filter(Site %in% c("UW-LakeShore", "UW-Sellery")),
    cases,
    by = c("Date", "Site"))
```

```{r}
write.csv(case_water, file = "data/uhs_housing_cases_water.csv", row.names = FALSE)
```

```{r}
case_waters <- case_water %>%
  select(Positive, Conductivity:Pct_BCoV, everything()) %>%
  pivot_longer(Positive:Pct_BCoV, names_to = "assay", values_to = "values") %>%
  filter(!is.na(values)) %>%
  mutate(wt = ifelse(assay == "AVG", wt, 1)) %>%
  arrange(Date) %>%
  group_by(Site, assay) %>%
  mutate(roll = roll7(values)) %>%
  ungroup
```

Now plot side by side cases and water meas over time, and consider time lags.

```{r fig.height = 10}
ggplot(case_waters %>% 
         filter(assay %in% c("Positive", "N1", "N2", 
                             "AVG","PMMoV","Pct_BCoV"))) +
  aes(Date, values, col = Site) +
  geom_smooth(aes(y = values), method = "lm", se = FALSE, size = 1,
              col = "black",
              data = case_waters %>% 
         filter(assay %in% c("Positive", "N1", "N2",
                             "AVG","PMMoV","Pct_BCoV"),
                Date >= max(Date) - 28 * 86400)) +
  geom_smooth(method = "loess", se = FALSE, size = 1, span = spanw) +
  geom_point(alpha = 0.5) +
  scale_y_log10() +
  facet_grid(assay ~ Site, scales = "free")
```


```{r}
ggplot(case_waters %>% 
         filter(assay %in% c("Positive", "AVG"),
                Site %in% c("UW-LakeShore", "UW-Sellery"))) +
  aes(Date, values, col = Site) +
  geom_smooth(aes(y = values), method = "lm", se = FALSE, size = 1,
              col = "black",
              data = case_waters %>% 
         filter(assay %in% c("Positive", "AVG"),
                Site %in% c("UW-LakeShore", "UW-Sellery"),
                Date >= max(Date) - 28 * 86400)) +
  geom_smooth(method = "loess", se = FALSE, size = 1, span = spanw) +
  geom_point(alpha = 0.5) +
  scale_y_log10() +
  facet_grid(assay ~ Site, scales = "free")
```
