---
title: "WasteWater Wisconsin"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4)
```

This document summarizes quarantine information. The R code file can be found in the [pandemic github repository](https://github.com/UW-Madison-DataScience/pandemic/blob/master/wastewaterWI.Rmd). The wastewater data is mostly at the city or smaller level; here it is aggregated to the county within HERC region. If a county has dramatic differences in levels across cities, that will go into the noise. There are also some issues with sites that don't correspond to cities I could identify (Red Cloud, Wolf) and some dates that were missing.

```{r}
library(tidyverse)
library(readxl)
library(broom)
library(RcppRoll)
kprint <- function(...) {
  if(interactive()) {
    print(...)
  } else {
    knitr::kable(...)
  }
}
```

```{r}
wicity <- read.csv("data/list-cities-wisconsin-200j.csv")
herc <- read.csv("data/herc.csv")
wwwi <- read_excel("data/wastewater/SARS-CoV-2 Project Data Compilation TO DHS w BCOV 201208.xlsx", 
                   col_types = c("text","date",rep("numeric",6)),
                   na = c("-","","0")) %>%
  rename(Date = "Collection Date",
         N1 = "N1: gc/1L",
         N2 = "N2: gc/1L",
         AVG = "Average gc/1L (N1 and N2)",
         PMMoV = "PMMoV: gc/1L",
         BCoV = "BCoV:  AVG % Recovery") %>%
  mutate(GEO = sqrt(N1 * N2),
         GEO = ifelse(is.na(N1), N2, GEO),
         GEO = ifelse(is.na(N2), N1, GEO)) %>%
  pivot_longer(N1:GEO, names_to = "assay", values_to = "value") %>%
  filter(!is.na(value),
         !is.na(Date)) %>%
  mutate(Location = ifelse(Location == "Jansewille", "Janesville", Location),
         Location = ifelse(Location == "Sun Praire", "Sun Prairie", Location),
         Location = ifelse(str_detect(Location, "MMSD"), "Madison", Location))
```

```{r}
wwwi <- wwwi %>%
  mutate(County = match(Location, wicity$City),
         County = ifelse(is.na(County), "Unknown", wicity$County.ies[County]),
         HERC = match(County, herc$county),
         HERC = ifelse(is.na(HERC), NA, herc$herc_region[HERC]))
```

```{r}
city_plot <- function(wwwi, city_name, assay_name = "GEO") {
  ggplot(wwwi %>% filter(Location == city_name, assay == assay_name)) +
      aes(Date, value) +
      scale_y_log10() +
      geom_point(alpha = 0.5, size = 2) +
      ggtitle(city_name) + 
      geom_smooth(method = "loess", se = FALSE, formula = "y ~ x")
}
```

```{r}
ggplot(wwwi) +
  aes(Date, value) +
  facet_grid(HERC ~ assay, scales = "free") +
  scale_y_log10() +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "loess", se = FALSE, formula = "y ~ x")
```

```{r}
ggplot(wwwi %>% 
         filter(assay == "AverageFlowRate")) +
  aes(Date, value, col = Location, group = Location) +
  facet_wrap(~ HERC, scales = "free_y") +
  scale_y_log10() +
  geom_point(alpha = 0.25) +
  theme(legend.position = "none") +
  geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
  ggtitle("Locations within HERC Region for AverageFlowRate")
```

```{r}
ggplot(wwwi %>% 
         filter(assay == "BCoV")) +
  aes(Date, value, col = Location, group = Location) +
  facet_wrap(~ HERC, scales = "free_y") +
  scale_y_log10() +
  geom_point(alpha = 0.25) +
  theme(legend.position = "none") +
  geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
  ggtitle("Locations within HERC Region for BCoV")
```

```{r}
ggplot(wwwi %>% 
         filter(assay == "PMMoV")) +
  aes(Date, value, col = Location, group = Location) +
  facet_wrap(~ HERC, scales = "free_y") +
  scale_y_log10() +
  geom_point(alpha = 0.25) +
  theme(legend.position = "none") +
  geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
  ggtitle("Locations within HERC Region for PMMoV")
```

```{r}
ggplot(wwwi %>% 
         filter(assay == "GEO",
                !is.na(HERC))) +
  aes(Date, value, col = Location, group = Location) +
  facet_wrap(~ HERC, scales = "free_y") +
  scale_y_log10() +
  geom_point(alpha = 0.25) +
  theme(legend.position = "none") +
  geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
  ggtitle("Locations within HERC Region for GEO")
```


```{r}
herc_plot <- function(wwwi, assay_name = "GEO", herc_name = "Fox Valley Area",
                      trend = 28) {
  if(is.na(herc_name)) {
    wwwi <- wwwi %>%
      mutate(HERC = ifelse(is.na(HERC), "Unknown", HERC))
    herc_name <- "Unknown"
  }
  p <- ggplot(wwwi %>%
                filter(assay == assay_name,
                       HERC == herc_name)) +
    aes(Date, value, col = Location, group = Location) +
    scale_y_log10() +
    geom_point(alpha = 0.5, size = 2) +
    ggtitle(paste("HERC", herc_name, " Assay ", assay_name))
  
#  if(trend > 0) {
    p <- p + 
      geom_smooth(aes(x = Date, y = value), method = "lm", se = FALSE, size = 2,
                  formula = "y ~ x",
                  data = wwwi %>%
                filter(assay == assay_name,
                       HERC == herc_name,
                       Date >= max(Date) - trend * 86400))
#  } else {
    p <- p +
      geom_smooth(aes(x = Date, y = value), method = "lm", se = FALSE, size = 2,
                  formula = "y ~ x",
                  data = wwwi %>%
                filter(assay == assay_name,
                       HERC == herc_name,
                       Date < max(Date) - trend * 86400))
#      geom_smooth(method = "loess", se = FALSE, formula = "y ~ x")
#  }
  p  
}
```

```{r}
herc_plot(wwwi, "GEO", "Fox Valley Area")
```

```{r}
herc_plot(wwwi, "GEO", "North Central")
```

```{r}
herc_plot(wwwi, "GEO", "Northeast")
```

```{r}
herc_plot(wwwi, "GEO", "Northwest")
```

```{r}
herc_plot(wwwi, "GEO", "South Central")
```

```{r}
herc_plot(wwwi, "GEO", "Southeast")
```

```{r}
herc_plot(wwwi, "GEO", "Western")
```

```{r}
herc_plot(wwwi, "GEO", NA)
```

#### Look at Janesville

```{r}
city_plot(wwwi, "Janesville") +
  ggtitle("Janesville GEO")
```

```{r}
city_plot(wwwi, "Janesville", "BCoV") +
  ggtitle("Janesville BCoV")
```

```{r}
dat <- wwwi %>%
  filter(Location == "Janesville", assay %in% c("GEO", "BCoV")) %>%
  select(Location, Date, assay, value) %>%
  pivot_wider(names_from = "assay", values_from = "value") %>%
  filter(!is.na(GEO),
         !is.na(BCoV),
         GEO > 1000)
fit <- lm(log10(GEO) ~ log10(BCoV), dat)
dat <- dat %>%
  mutate(resid = resid(fit),
         resid = 10^(mean(log10(GEO)) + resid))
```

```{r}
ggplot(dat) +
      aes(Date, resid) +
      scale_y_log10() +
      geom_point(alpha = 0.5, size = 2) +
      ggtitle("Janesville") + 
      geom_smooth(method = "loess", se = FALSE, formula = "y ~ x") +
  ggtitle("Janesville GEO remove low value and regressed on BCoV")
```

#### Look at individual cities

Analysis above may be flawed by aggregating disparate cities or locations within a county. Just a beginning look at cities or other locations.

```{r}
ggplot(wwwi %>% filter(assay == "GEO", !is.na(HERC))) +
  aes(log10(value), col = Location) +
  geom_density() +
  facet_wrap(~ HERC) +
  theme(legend.position = "none") +
  xlab("log10(GEO)")
```

```{r}
dat <- wwwi %>%
  filter(assay %in% c("GEO","PMMoV","BCoV","AverageFlowRate")) %>%
  select(Location, assay, value, Date) %>%
  # 2020-11-17 has two entries for Wausau
  distinct(Location, assay, Date, .keep_all = TRUE) %>%
  pivot_wider(names_from = "assay", values_from = "value") %>%
  filter(!is.na(GEO),
         GEO > 1000)
fit <- lm(log10(GEO) ~ log10(PMMoV) + log10(BCoV) + AverageFlowRate + Date * Location, dat)
drop1(fit, fit, test = "F")
```

```{r}
city_plot(wwwi, "Appleton")
```

```{r}
ggplot(dat %>% filter(Location == "Janesville", GEO > 1000)) +
  aes(GEO, BCoV, col = Date) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
  scale_x_log10() +
  scale_y_log10()
```

```{r}
city_plot(wwwi, "Janesville")
```

```{r}
city_plot(wwwi, "Kenosha")
```

```{r}
city_plot(wwwi, "Madison")
```

```{r}
city_plot(wwwi, "Manitowoc")
```

```{r}
city_plot(wwwi, "Marshfield")
```

```{r}
city_plot(wwwi, "Oshkosh")
```

```{r}
city_plot(wwwi, "Platteville")
```

```{r}
city_plot(wwwi, "River Falls")
```

```{r}
city_plot(wwwi, "Sheboygan")
```

```{r}
city_plot(wwwi, "Stevens Point")
```

```{r}
city_plot(wwwi, "Sun Prairie")
```

```{r}
city_plot(wwwi, "Viroqua")
```

```{r}
city_plot(wwwi, "Whitewater")
```


